---
title: "How Often Does it Rain in Vancouver?"
Author: "Asen Lee"
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    number_sections: false
    toc: FALSE
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, error = FALSE, echo = TRUE)
library(tidyverse)
library(ggplot2)
library(caret)
```

# Introduction

Vancouver is known to be a rainy city, and as a student attending UBC, it is only natural for us to wonder what proportion of days it actually rains, or the average number of rainy days per year. Using different sampling methods, I will conduct an analysis on Vancouver weather data over the past 83 years to estimate the total proportion of rainy days in Vancouver’s history, and the mean number of days per year where it rains over 10mm. I will use both SRS and Stratified Sampling as our methods to draw from a dataset of daily Vancouver International Airport weather data from 1937–2020.

## Background

Environment Canada defines a rainy day as a calendar day with a total rainfall of at least 0.2mm. Note that rainfall is differentiated from precipitation, as the latter includes snowfall. Summarized weather data from Environment Canada Canadian Climate Normals 1981–2010 Station Data shows that average annual number of days with rainfall is 165.1, a proportion of 45.23%. We treat this statistic as a point of reference. We cannot treat this as a true population parameter however, as this statistic arose from data between 1981–2010. The City of Vancouver was incorporated in 1886, thus our parameter of interest is the proportion of rainy days in all of Vancouver’s existence. There is no reliable weather data extending to that time, therefore the true population parameter is unknown.

## Objectives

We denote our two parameters of interest, π : the proportion of rainy days in Vancouver, and µ : the mean number of days per year where it rains at least 10mm. The population of interest are all calendar days in Vancouver’s history. We choose 10mm as a “cut off” for our second estimate, as we consider daily rainfall amounts of less than 10mm as light rain. Hence, we are interested not only in the proportion of rainy days, but also in how many of those days are not considered “light rain” days.


# Methods

## Data Collection

We obtained daily weather records at Vancouver International Airport from Environment Canada for years 1937–2020. Each year of weather data are available for download as a .csv file on Environment Canada’s website. We combined all 83 years of .csv files into a single .csv dataset of 30,681 days of weather records. For each daily record, the dataset contains information such as low, high and mean temperatures, total rain in mm, total snowfall in cm, and total precipitation in mm. After dataset cleaning, we only retain the fields: date, mean temperature and total rain.


```{r}
data <- read.csv("combined_weather.csv")
```

```{r, echo = FALSE}
# data wrangling
data <- data[data$Year != "Year", ]
data <- data %>% select(Year, Month, Day, Mean.Temp..Â.C., Total.Rain..mm.)
data <- data %>% rename(mean_temp = Mean.Temp..Â.C.,
                        total_rain = Total.Rain..mm.)
data[] <- lapply(data, function(x) {
   if (is.character(x))
        as.numeric(x)
   else x
})
data <- data %>% fill(total_rain)
data <- data %>% fill(mean_temp)
```


## Assumptions

We assumed that the proportion of rainy days each year are independent and identically distributed. Hence by Central Limit Theorem, our estimators for π and µ are reasonable and we can extend our inferences to dates outside of the dataset.


## Sample Size Calculation:
\
Our population of interest are all calendar days in Vancouver’s history, which are all days from 1886–present. In theory, population size N, is very large and is increasing each day, so $N \rightarrow \infty$. However, the earliest reliable data we were able to obtain is from 1937, so we are sampling from a finite dataset of $N = 30681$. How we are able to extend our inferences to dates outside of this dataset is explained in Assumptions. Our sample size n was calculated with forumla: $n = n_0/(1 + n_0/N)$ where $n_0 = 1.96^2s_g^2/\delta^2$, $s_g = 0.5$ (conservative guess) and $\delta = 0.01, 0.02, 0.03$ to calculate the sample size. $\delta$ is the margin of error at 95% confidence and $s_g$ is our prior guess of the variance of the parameter (worst case).

```{r}
# sample size calculation
N <- nrow(data)
s_g <- 0.5
delta <- c(0.03, 0.02, 0.01)
n_0 <- (1.96^2*s_g^2)/delta^2
n <- n_0/(1 + n_0/N)

# table
mat <- matrix(c(delta, n), nrow = 3)
knitr::kable(x = mat, col.names = c("delta", "n"))
```
For our analysis, we want an accuracy of +/- 0.02 percentage points 19 times out of 20. Hence, we will choose our sample size to be 2227.
\
We will collect a simple random sample of size $n = 2227$ from a total population of $N = 30681$ with replacement to estimate the proportion of rainy days in Vancouver.

## Simple Random Sampling

```{r}
# population proportion
set.seed(2021)
n <- 2227
srs <- sample(data$total_rain, size = n, replace = FALSE)
p_hat <- length(srs[srs > 0])/n
round(p_hat, 4)
```
The population proportion p is found to be 0.4486.\

We will use the formula: $SE(\hat{p}) = \sqrt{\hat{p}(1-\hat{p})/n}$ to calculate the standard error.

```{r}
# standard error
se <- sqrt(p_hat*(1-p_hat)/n)
round(se, 4)
```
We get a standard error of 0.105, which is expected since we chose the sample size to give +/- 0.02 percentage points.
\
95% Confidence Interval: \
```{r}
CI <- c(p_hat - 1.96*se, p_hat + 1.96*se)
round(CI, 4)
```
We obtain a 95% confidence interval of [0.4279, 0.4692]. Hence, we are 95% confident that out of all the days in Vancouver, 0.4279 to 0.4692 percent of days will be rainy days. This is quite surprising although we knew that it rains often in Vancouver.\
This also means that out of 365 days, we have about 161 rainy days!
\
The estimate for $µ$, mean number of days per year in Vancouver where it rains at least 10mm is:

```{r}
# mu
srs <- sample(data$total_rain, size = n, replace = FALSE)
srs <- as.numeric(as.character(srs))
p_hat <- length(srs[srs >= 10])/n
mu_hat <- p_hat*365
round(mu_hat,0)
```

with standard error of 2.3349.

```{r}
# SE mu
SE.mu <- sqrt(365^2*(1-n/N)*p_hat*(1-p_hat)/n)
round(SE.mu,4)
# 95 CI
CI <- c(mu_hat - 1.96*SE.mu, mu_hat + 1.96*SE.mu)
round(CI, 0)
```

Hence, the 95% confidence interval for µ is (36, 45).

## Stratified Sampling
\
We divide the dataset into four equal size strata based on season: months 12, 1, 2
as winter, 3–5 as spring, 6–8 as summer, and 9–11 as fall.
\
### Within-stratum Variance

As we have lived in Vancouver all our lives, we believe that the variance of daily weather may differ between seasons. For example, summers tend to be dry and warm with minimal variability between each day, whereas some winters can be very wet and mild while other winters dry and cold. We therefore attempted to take this variability into account when determining the sample size’s of our strata.

From Environment Canada, we obtained a dataset outlining annual total rainfall amounts in mm, by season, from years 1897–2012.

With this data, we computed sample variances of the total rainfall amounts for each season. The results are as follows:


```{r}
# estimated variance of individual stratum
seas.rainfall <- read.csv("van_adjusted_rainfall.txt", header = T )
seas.rainfall <- seas.rainfall %>% select(Winter, Spring, Summer, Autumn)
seas.rainfall.var <- apply(seas.rainfall, MARGIN = 2, var)
# table
mat <- matrix(c(12039, 5017, 2297, 8937), nrow = 1)
knitr::kable(x = mat, col.names = c("Winter", "Spring", "Summer", "Fall"))
```

Note that these are the variance of total annual rainfall amounts, which is not the same as the variance of the proportion of rainy days. These variances were used relative to each other to compute stratum sample sizes with optimal allocation. An important assumption we make is that seasons with greater variance in total rainfall each year also have a greater variance in the probability of a given day to be rainy.

### Stratum Sample Sizes
\
We used optimal allocation to determine the sample sizes for each stratum. Denote stratum h = 1, 2, 3, 4 as Winter, Spring, Summer, and Fall respectively. Assume all years are equally divided into four seasons so stratum population size N1 = N2 = N3 = N4.

```{r}
# Compute sample size by optimal allocation
n <- 2227 #total sample size
Nh <- 1/4 #assume the year is divided up into four equal seasons
nh <- c(Nh*sqrt(seas.rainfall.var[1]), Nh*sqrt(seas.rainfall.var[2]), 
        Nh*sqrt(seas.rainfall.var[3]), Nh*sqrt(seas.rainfall.var[4]))
nh <- c( nh[1] / sum(nh), nh[2] / sum(nh), nh[3] / sum(nh), nh[4]/sum(nh) )*n
nh <- round(nh,0)
nh
```

```{r}
# Divide dataset into four strata based on season
# winter = 12,1,2
# spring = 3,4,5
# summer = 6,7,8
# fall = 9,10,11
winter <- data %>% filter(Month=="01" | Month=="1" | Month==1 | 
                          Month=="02" | Month=="2" | Month==2 | 
                          Month=="12" | Month==12)
spring <- data %>% filter(Month=="03" | Month=="3" | Month==3 |
                          Month=="04" | Month=="4" | Month==4 |
                          Month=="05" | Month=="5" | Month==5)
summer <- data %>% filter(Month=="06" | Month=="6" | Month==6 |
                          Month=="07" | Month=="7" | Month==7 |
                          Month=="08" | Month=="8" | Month==8)
fall <- data %>% filter(Month=="09" | Month=="9" | Month==9 |
                        Month=="10" | Month==10 |
                        Month=="11" | Month==11)
```


### Estimation
\
Once again, our first parameter of interest is the proportion of days where total rain ≥ 0.2mm, denoted $π_s$. After we obtained a simple random sample of size $n_h$ for each stratum h, denote πh as the proportion of rainy days in stratum h, $y_hi$ be the value of total rain on the ith day in the sample of stratum h, and $u_hi$ = 1 if $y_hi$ ≥ 0.2, else $u_hi$ = 0.
\
Firstly, we obtained estimates for $π_h$, proportion of rainy days in Vancouver in each season h.

```{r}
# Create samples
set.seed(2021)
winter.str <- sample(winter$total_rain, size = nh[1], replace = FALSE)
winter.str <- as.numeric(as.character(winter.str))
spring.str <- sample(spring$total_rain, size = nh[2], replace = FALSE)
spring.str <- as.numeric(as.character(spring.str))
summer.str <- sample(summer$total_rain, size = nh[3], replace = FALSE)
summer.str <- as.numeric(as.character(summer.str))
fall.str <- sample(fall$total_rain, size = nh[4], replace = FALSE)
fall.str <- as.numeric(as.character(fall.str))
# Estimate str proportion of rainy days
p.winter <- length(winter.str[winter.str >= 0.2])/nh[1]
p.spring <- length(spring.str[spring.str >= 0.2])/nh[2]
p.summer <- length(summer.str[summer.str >= 0.2])/nh[3]
p.fall <- length(fall.str[fall.str >= 0.2])/nh[4]
p.h <- c(p.winter, p.spring, p.summer, p.fall)
round(p.h, 4)
```
\
The corresponding standard errors for each $π_h$ are:
```{r}
# SE_h
se.h <- c(sqrt((1-nh[1]/nrow(winter))*(p.h[1]*(1-p.h[1]))/nh[1]),
sqrt((1-nh[2]/nrow(spring))*(p.h[2]*(1-p.h[2]))/nh[2]),
sqrt((1-nh[3]/nrow(summer))*(p.h[3]*(1-p.h[3]))/nh[3]),
sqrt((1-nh[4]/nrow(fall))*(p.h[4]*(1-p.h[4]))/nh[4]))
round(se.h, 4)
```
\
We then combined them to form a single estimate for $π_s$ across all strata (all seasons).

```{r}
p.str <- 1/4*sum(p.h)
round(p.str, 4)
```
With Standard error of 0.0104.

```{r}
# SE str
se.str <- sqrt((1/4)^2*sum(se.h^2))
round(se.str, 4)
```
\
Hence, the 95% confidence interval for $π_s$ is (0.428, 0.468).

```{r}
# 95% CI
CI <- c(p.str-1.96*se.str, p.str+1.96*se.str)
round(CI, 4)
```

Our second parameter of interest $µ_s$ is the mean number of rainy days per year across strata where total rain ≥ 10mm. Let $v_hi$ = 1 if $y_hi$ ≥ 10, else $v_hi$ = 0. 

The means and the standard errors for the means are listed below:

```{r}
# mu
mu.winter <- length(winter.str[winter.str >= 10])/nh[1]
mu.spring <- length(spring.str[spring.str >= 10])/nh[2]
mu.summer <- length(summer.str[summer.str >= 10])/nh[3]
mu.fall <- length(fall.str[fall.str >= 10])/nh[4]
mu.h <- c(mu.winter, mu.spring, mu.summer, mu.fall)*(365/4)
round(mu.h,0)
# SE
se.h <- c(sqrt((365/4)^2*(1-nh[1]/nrow(winter))*(mu.winter*(1-mu.winter))/nh[1]),
sqrt((365/4)^2*(1-nh[2]/nrow(spring))*(mu.spring*(1-mu.spring))/nh[2]),
sqrt((365/4)^2*(1-nh[3]/nrow(summer))*(mu.summer*(1-mu.summer))/nh[3]),
sqrt((365/4)^2*(1-nh[4]/nrow(fall))*(mu.fall*(1-mu.fall))/nh[4]))
round(se.h, 4)
```

After combining the estimates across strata, we get our estimate for mean number of days per year (and its SE) where it rains at least 10mm in Vancouver:

```{r}
# mu_str
mu.str <- sum(mu.h)
round(mu.str,0)
# SE str
se.str <- sqrt(sum(se.h^2))
round(se.str, 4)
```
Hence, the 95% confidence interval is (32,40).

```{r}
# 95% CI
CI <- c(mu.str-1.96*se.str, mu.str+1.96*se.str)
round(CI, 0)
```

# Conclusion
\
Firstly, with our estimate of $π$ we see that they are both very similar and there is a negligible difference between the standard errors. Since there is no decrease in margin of error, there is no improvement in the stratified sampling estimate. Similarly, for our estimate of µ, we see that the two confidence intervals have the same width and the difference in standard error is negligible. Therefore there is also no improvement in stratified sampling over simple random sample

